Invasive lionfish present region-wide variation of allometric growth in the Western Atlantic

### 2018-02-25 Manuscript.tex
Added new version of manuscript

\title{Allometric growth parameters for invasive lionfish \emph{Pterois
volitans} (Actinopterygii, Scorpaenidae) in the Western Atlantic}

@@

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, inform lionfish removal programs, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass, frequently from length observations. This work compares
length-weight relationships of the invasive lionfish through the
invasion range and reports the length--weight relationship for lionfish
in the Central Mexican Caribbean. A review of 13 length--weight
relationships reported in eight peer-reviewed studies and FishBase is
provided. These parameters were used to identify spatial variation in
weight-at-length. For a given length, parameters from the Caribbean
yielded lower weights than those from the Gulf of Mexico and Atlantic,
indicating that weight-at-length is spatially variable. This highlights
the importance of using site-specific parameters to estimate biomass
from length observations. This study also reports a new pair of
length-weight parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for
organisms sampled in the Central Mexican Caribbean. Findings from this
work can aid managers and decision makers to better select length-weight
parameters when these are not available for their region of interest.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

@@

At least 84\% of the marine eco-regions have reported the presence of an
invasive species \citep{molnar_2008}. These represent a major threat to
local biodiversity and the economic activities that depend on it, like
tourism or fisheries \citep{bax_2003}. Invasive species may also
threaten native species through competition \citep{davis_2003} or
predation. By 2005, the economic cost of invasive species to the United
States was estimated at \$120 billion per year and nearly 42\% of
species that have been included in the Endangered or Threatened species
list have been labeled as such due to presence of invasive species
\citep{pimentel_2005}. This highlights the importance of understanding,
managing, and preventing ecological invasions.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the North-Western Atlantic and the Caribbean, likely introduced
through liberation of aquarium-kept organisms \citep{betancurr_2011}.
They are the first marine vertebrates to establish in North Atlantic
\citep{schofield_2009,schofield_2010} and Caribbean coasts
\citep{sabidoitza_2016}. Lionfish have been widely reported in coral
reefs \citep{aguilarperera_2010}, but also in other habitats such as
estuaries \citep{jud_2011}, mangroves \citep{barbour_2010}, areas with
hard-bottoms \citep{muoz_2011}, and mesophotic reefs
\citep{andradibrown_2017}. Due to its threat to local biodiversity, the
speed of their spread, and its difficulty of management, their presence
in these waters has been labeled as a major marine invasion
\citep{hixon_2016}.

A significant amount of research has been done to describe lionfish
feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas
\citep{morris_2009,cote_2013}, Northern Gulf of Mexico
\citep{dahl_2014}, Mexican Caribbean
\citep{valdezmoreno_2012,villaseorderbez_2014}, Belize
\citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}. Their
feeding behavior and high consumption rates can reduce recruitment
\citep{albins_2008} and population sizes \citep{green_2012} of native
reef-fish species, and further the endangerment of critically endangered
reef fish \citep{rocha_2015}. (However, see \citet{hackerott_2017} for a
case where there was no evidence that lionfish affected the density,
richness, or composition of prey fishes). Major efforts have also been
made to understand the possible impacts of the invasion by keeping track
of its range through time \citep{schofield_2009,schofield_2010} and
predicting invasion ranges under climate change scenarios
\citep{grieve_2016}. By combining information from these disciplines,
researchers have been able to predict the trophic impacts of lionfish
\citep{ariasgonzalez_2011}, which can then be translated into
ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption \citep{chin_2016}. In some cases, these have shown to
significantly reduce -but not quite eliminate- lionfish abundances at
local scales \citep{sandel_2015,chin_2016,deleon_2013}. The rapid
recovery rates exhibited by lionfish \citep{barbour_2011} and the
persistent populations in mesophotic coral ecosystems
\citep{andradibrown_2017} -which can contribute with recruitment to
shallow-water populations- make of complete eradication through fishing
effort an unlikely solution. However, further incentivizing its
consumption might create a demand big enough to promote and sustain a
stable fishery \citep{chin_2016}, which can reduce local abundances and
control -not eradicate- the invasion while providing alternative
livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field
observations\citep{usseglio_2017,sandel_2015,chin_2016,deleon_2013} and
empirical modeling \citep{barbour_2011,morris_2011,johnston_2015}. The
latter measure changes in biomass or density
\citep{barbour_2011,johnston_2015} in response to increased mortality
(\emph{i.e.} lionfish removal). In this case, biomass represents the sum
of all fish's individual weight. Total Weight (TW) can be estimated from
Total Length (TL) observations using the allometric growth equation
(\(TW = aTL^b\)). Parameters \(a\) and \(b\) for this equation exist for
North Carolina \citep{barbour_2011}, Northern \citep{fogg_2013} and
Southern Gulf of Mexico \citep{aguilarperera_2016}, the Southern Mexican
Caribbean \citep{sabidoitza_2016}, Little Cayman \citep{edwards_2014},
Jamaica \citep{chin_2016}, Bonaire \citep{deleon_2013} and Costa Rica
\citep{sandel_2015}, but remain unavailable for the central Mexican
Caribbean. The weight-at-length of a species can vary across regions as
a response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}. Thus,
when using biomass-informed models or estimating biomass from length
observations, it is important to use site-specific parameters to obtain
an accurate estimate. This is especially important when research
involves identifying the total biomass available for harvest by fishers
\citep{chin_2016} or the efficacy of lionfish removals
\citep{barbour_2011,morris_2011,johnston_2015}.

Here, I provide the first allometric growth parameters for the invasive
lionfish in the central Mexican Caribbean region. At the same time, I
highlight the importance of using site-specific parameters by estimating
biomass with parameters from other regions across the invasion range and
comparing them to observed biomass. I also provide other 13 standardized
parameters from eight studies through the invasion range, making them
readily accessible for future research. Finally, I discuss the way in
which allometric parameters are reported, and call for standardization
to facilitate their use.

@@

\emph{Area of study}

The study took place off the coasts of Playa del Carmen, in the Mexican
Caribbean Figure \ref{fig:map}. The region represents the northernmost
section of the Mesoamerican Barrier Reef System \citep{ruizzarate_2004}.
Coral reefs and mangroves are locally important habitats that represent
important sources of income in terms of extractive (\emph{e.g.}
recreational fishing) and non-extractive (\emph{e.g.} SCUBA diving)
activities related to tourism, the main source of income to the local
economy \citep{murray_2007}.

The reef profile has been described by \citet{ariasgonzalez_1998},
indicating that the reef lagoon extends about 500 m from the coast,
until the reef crest is reached. The reef becomes deeper, leading to the
reef front often found at 700 m from the coastline and extends for an
additional 300 m. At approximately 1000 m away from shore and 30 - 40 m
depth, the reef leads to a drop-off. Along a perpendicular profile to
the coast, bands of reef are interrupted by sand patches at 8 - 12 m
deep and 16-18 m deep. Along the coast, these reefs have been reported
to be under significant anthropogenic pressure, likely causing a shift
in structure and function \citep{bozec_2008}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Map of the study area.
The black inset on the left (Yucatan Peninsula) indicates the location
where study sites are distributed. On the right, circular markers
indicate sampling sites and the black romboid indicates location of
Puerto Aventuras, Mexico.}
\end{figure}

\emph{Fish sampling}

A total of 33 SCUBA immersions were performed in 10 sampling sites along
the coast in 2010 (Fig. 1, Table I). Sampling locations included wall
and carpet reefs at depths between 5.7 m and 38.1 m. All observed
organisms (n = 109) were collected using hand nets and numbered
collection bottles. The use of hand nets prevented any weight loss due
to bleeding and allowed a better representation of small sizes, often
ignored due to gear selectivity when spearing. Organisms were euthanized
and frozen within 30 minutes of completing the dive and stored for
posterior analysis. Total Length (TL; mm) and Total Weight (TW; gr) were
recorded for all organisms.

\caption{(\#tab:table of locations)Coordinates, minimum, maximum and mean depth (m), and number of samples for each location.}
\centering
\begin{tabular}[t]{llllllr}
\toprule
Location & Lat. & Long. & Min. Depth & Max. Depth & Mean Depth & n\\
\midrule
Canones & 20.477 & -87.233 & 15.0 & 31.2 & 21.6 & 11\\
Castillo & 20.496 & -87.220 & 12.5 & 30.5 & 27.5 & 18\\
Cuevitas & 20.478 & -87.244 & 7.4 & 12.8 & 11.2 & 4\\
Islas & 20.490 & -87.228 & 14.0 & 19.4 & 16.7 & 10\\
Paamul & 20.513 & -87.192 & 9.9 & 22.7 & 15.5 & 31\\
\addlinespace
Paraiso & 20.484 & -87.226 & 9.4 & 38.1 & 17.7 & 16\\
Pared & 20.502 & -87.212 & 12.1 & 21.0 & 16.3 & 12\\
Pedregal & 20.507 & -87.204 & 14.4 & 14.9 & 14.7 & 3\\
Santos & 20.493 & -87.222 & 5.7 & 26.6 & 16.2 & 2\\
Tzimin-Ha & 20.393 & -87.307 & 21.2 & 24.6 & 22.9 & 2\\
Total &  &  & 5.7 & 38.1 & 18.6 & 109\\
\bottomrule
\end{tabular}
\end{table}

\emph{Data analysis}

The weight at length relationship between the observed variables is
described by the allometric growth function:

@@

Where \(Y = log_{10}(TW)\), \(X = log_{10}(TL)\), \(m = b\), and
\(c = log_{10}(a)\). Since \(b = m\), we will only use \(b\) throughout
the paper for simplicity. The coefficients (\(c\) and \(b\)) were
estimated with an Ordinary Least Square Regression and
heteroskedastic-robust standard error correction was performed. Both
coefficients were tested against the null hypothesis of no change
(\emph{i.e.} \(H_0: c = 0\) and \(H_0: b = 0\)). Additionally, the
allometric parameter was tested against the null hypothesis of isometric
growth (\(H_0: b = 3\)). Coefficients were tested with a two-tailed
Student's t-test. The significance of the regression was corroborated
with an F-test.

Other length-weight relationships (n = 13) were extracted from
peer-reviewed literature. Parameters were obtained for North Carolina
{[}n = 1,barbour\_2011{]}, Northern {[}n = 3,fogg\_2013{]} and Southern
Gulf of Mexico {[}n = 3,aguilarperera\_2016{]}, the Southern Mexican
Caribbean {[}n = 1,sabidoitza\_2016{]}, Little Cayman {[}n =
1,edwards\_2014{]}, Jamaica {[}n = 1,chin\_2016{]}, Bonaire {[}n =
1,deleon\_2013{]} and Costa Rica {[}n = 1,sandel\_2015{]} and Fishbase
{[}n = 1,froese\_website\_2016{]} were also included. When available,
information on sampling methods, gender differentiation, location, and
depth ranges of each study was retrieved. Whenever gender was not
specified, it was assumed that the results were presented for pooled
genders.

During the review process, some papers indistinctly used \(a\) to report
either the ponderal index in eq. 1 or the y-intercept (\(c\)) in eq. 3,
which might sometimes be overlooked. Furthermore, some studies reported
their parameters as mm-to-gr conversions, but a rapid evaluation of such
parameters indicated that they were estimated as cm-to-gr conversions.
Here, all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

Since uncertainty arround estimated relationships was not reported in
some of the reviewed studies, it was not possible to test for
statistical differences between relationships. Instead, the 13
length-weight relationships were usd to calculate expected weight for
the organisms sampled in the Central Mexican Caribbean (n = 109).
Expected weights were divided by the observed weights to obtain a ratio.
Difference in mean weight ratios across studies were tested with a
one-way Analysis of Variance (ANOVA).

All hypothesis testing was performed with an \emph{a priori} confidence
level of \(\alpha = 0.01\) in R version 3.4.0 \citep{rcore_2017}. Data
wrangling was done with the tidyverse package \citep{tidyverse_2017}.
Maps were created with a mix of functions from the sp \citep{sp_2017},
rgdal \citep{rgdal_2017}, tmap \citep{tmap_2017}, and tmaptools
\citep{tmaptools_2017} packages. Heteroskedastic-robust standard errors
were calculated with the sandwich \citep{sandwich_2014} and lmtest
\citep{lmtest_2002} packages. Models were manipulated with the broom
package \citep{broom_2017}. Raw data and code used in this work is
available at github.com/jcvdav/lionfish\_biometry.

@@

Organism TL ranged between 34 and 310 mm and TW between 0.3 and 397.7
gr. The smallest organism (TL = 34.00 mm) was also the lightest organism
(TW = 0.30 gr). However, the largest organism (TL = 310.00 mm) was not
the heaviest (TW = 303.70 gr), and the heaviest organism (TW = 397.70
gr) was 292.00 mm in total length. Kernell density plots (Fig. 2) show
the distribution for TL and TW for all sampled organisms. Both measures
were positively skewed, with skewness of 0.87 for TL and 2.25 for TW.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/histogram figures-1.pdf}
\caption{(\#fig:histogram figures)Kernell density plots for a) Total
length (mm) and b) Total weight (gr) for 109 lionfish sampled in the
central Mexican Caribbean.}
\end{figure}

\emph{Length-weight relationship}

The model adjusted to eq. 3 estimated the coefficient values at
\(b = 3.2347391\) and \(c = -5.4940866\). Thus, TW (gr) can be
calculated from TL (mm) as a linear equation:
\(log_{10}(TW) = 3.2347391\times log_{10}(TL) -5.4940866\), or its
exponential form: \(TW = 3.2056297\times 10^{-6}\times TL^{3.2347391}\).
The intercept (\(c\)) and slope \((b)\) were significantly different
from zero (\(t(107) = -66.17; p<0.01\) and \(t(107) = 83.24; p<0.01\),
respectively), rejecting the null hypothesis of no change. Additionally,
the allometric factor (\(b\)) was significantly different from the value
of isometric growth of \(b = 3\) (\(t(107) = 6.04; p<0.01\)), indicating
that lionfish present allometric growth. More information on model fit
and confidence intervals for the estimated coefficients is presented in
Table II. The relationship between Total Length and Total Weight is
presented in Figure 3.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/scatter plot-1.pdf}
\caption{(\#fig:scatter plot)Length-weight relationship for 109 lionfish
sampled in the central Mexican Caribbean. Points indicate samples, solid
line indicates curve of best fit.}
\end{figure}

\begin{table}[!htbp] \centering 
  \caption{Regression table for the linear model fit between log10-transformed Total Weight (dependent variable) and Total Length (independent variable). Numbers in parentheses next to coefficient estimates indicate heteroskedastic-robust standard errors.} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & log10(TW) \\ 
\hline \\[-1.8ex] 
 c & $-$5.494 (0.083)$^{***}$ \\ 
  b & 3.235 (0.039)$^{***}$ \\ 
 \hline \\[-1.8ex] 
95\% CI for c & (-5.657--5.331) \\ 
85\% CI for b & (3.159-3.311) \\ 
F Statistic & 6928.67*** (df = 1; 107) \\ 
Observations & 109 \\ 
Adjusted R$^{2}$ & 0.976 \\ 
Residual Std. Error & 0.096 (df = 107) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\emph{Comparison of allometric parameters}

From the eight peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and Fishbase
\citep{froese_website_2016}, 13 parameters were identified (Table III).
Two studies \citep{aguilarperera_2016,fogg_2013} reported gender-level
and pooled parameters, while the rest presented pooled results. The
smallest coefficient of determination was presented by \citet{chin_2016}
with \(R^2 = 0.8715\), while \citet{sabidoitza_2016} reported the
highest value at \(R^2 = 0.9907\). Reviewed studies presented
information for organisms obtained at depths between 0.5 and 57 m. Two
studies \citep{aguilarperera_2016,chin_2016} explicitly stated that
their organisms were sampled with pole spears. Five studies
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016}
mentioned that some of their organisms were obtained with pole spears
(or other type of harpoon). A single study \citep{deleon_2013} did not
specify how organisms were sampled.

Parameters from models fit to males or females exclusively tend to have
a higher steepness (\emph{i.e.} higher allometric parameter), with mean
\(\pm\) standard deviation values of \(b = 3.27 \pm 0.06\) and
\(b = 3.31 \pm 0.23\) for males and females respectively, compared to
parameters from models for pooled genders with a mean \(\pm\) standard
deviation value of \(b = 3.09 \pm 0.22\). In the case of the ponderal
index (\(a\)) and its \(log_{10}\) transformed parameter (\(c\)), values
were higher for parameters for pooled genders. Figure 4 shows the
predicted weights for organisms within the size range of these study
using the 14 parameters previously described.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/review plots-1.pdf}
\caption{(\#fig:review plots)Length-weight relationships (n = 14) for
eight studies, this study, and Fishbase. Colors indicate studies from
which the parameters were extracted. Solid lines indicate that the fit
was performed for males and females pooled together. Dotted lines
indicate that the regression was performed on females, and dashed lines
indicate it was performed for males.}
\end{figure}

There were significant differences in expected-to-observed weight ratios
estimated for each pair of parameters (F(13, 1512) = 39.28; p
\textless{} 0.05). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
expected-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,barbour\_2011 yielded the highest weight
estimates, with a mean (\(\pm\) SD) expected-to-observed weight ratio of
1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure 5.

>>>>

\title{Invasive lionfish present region-wide variation of allometric growth in
the Western Atlantic}

@@

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, inform lionfish removal programs, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass, frequently from length observations. This work compares
length-weight relationships of the invasive lionfish through the
invasion with the addition of parameters for the Central Mexican
Caribbean. A review of 15 length-weight relationships reported in 10
peer-reviewed studies shows that lionfish exhibit spatial variation in
weight-at-length. The reviewed parameters indicate that, for the same
length, lionfish in the Caribbean have lower body mass than in the
Atlantic or Gulf of Mexico. This highlights the importance of using
site-specific parameters to estimate biomass from length observations.
This study also reports a new pair of length-weight parameters
(\(a = 3.2056\times 10^{-6}; b = 3.235\)) for organisms sampled in the
Central Mexican Caribbean. Findings from this work can aid managers and
decision makers to better select length-weight parameters when these are
not available for their region of interest.
\end{abstract}

\begin{document}

@@

At least 84\% of the marine eco-regions have reported the presence of an
invasive species \citep{molnar_2008}, which represent a major threat to
local biodiversity and the economic activities that depend on it
\citep{bax_2003}. Invasive species may threaten native species through
predation, competition, or indirect habitat effects
\citep{davis_2003, gurevitch_2004}. By 2005, the economic cost of
invasive species to the United States was estimated at USD\$120 billion
per year\citep{pimentel_2005}.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic and the Caribbean, likely introduced through
liberation of aquarium-kept organisms \citep{betancurr_2011}. They are
the first marine vertebrates to establish in North Atlantic
\citep{schofield_2009,schofield_2010} and Caribbean coasts
\citep{sabidoitza_2016}. Lionfish have been widely reported in coral
reefs \citep{aguilarperera_2010}, but also in other habitats such as
estuaries \citep{jud_2011}, mangroves \citep{barbour_2010}, areas with
hard-bottoms \citep{muoz_2011}, and mesophotic reefs
\citep{andradibrown_2017}. Due to their threat to local biodiversity,
the speed of their spread, and difficulty of management, their presence
in these waters has been labeled as a major marine invasion
\citep{hixon_2016}.

A significant amount of research has been done to describe lionfish
feeding ecology in North Carolina \citep{muoz_2011}, the Bahamas
\citep{morris_2009,cote_2013}, Northern Gulf of Mexico
\citep{dahl_2014}, Mexican Caribbean
\citep{valdezmoreno_2012,villaseorderbez_2014}, Belize
\citep{hackerott_2017}, and Costa Rica \citep{sandel_2015}. Their
feeding behavior and high consumption rates can reduce recruitment
\citep{albins_2008} and population sizes \citep{green_2012} of native
reef-fish species, and further the endangerment of critically endangered
reef fish \citep{rocha_2015}. (However, see \citet{hackerott_2017} for a
case where there was no evidence that lionfish affected the density,
richness, or community composition of prey fishes). Major efforts have
been made to understand the possible impacts of the invasion by keeping
track of its range through time \citep{schofield_2009,schofield_2010}
and predicting invasion ranges under future climates
\citep{grieve_2016}. By combining information from these disciplines,
researchers have been able to predict the trophic impacts of lionfish
\citep{ariasgonzalez_2011}, which can then be translated into
ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption \citep{chin_2016}. In some cases, these have shown to
significantly reduce --but not quite eliminate-- lionfish abundances at
local scales \citep{sandel_2015,chin_2016,deleon_2013}. The rapid
recovery rates exhibited by lionfish \citep{barbour_2011} and the
persistent populations in mesophotic coral ecosystems
\citep{andradibrown_2017} --which can contribute with recruitment to
shallow-water populations-- make of complete eradication through fishing
effort an unlikely solution. However, further incentivizing its
consumption might create a demand big enough to promote and sustain a
stable fishery \citep{chin_2016}, which can reduce local abundances and
control the invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field
observations\citep{usseglio_2017,sandel_2015,chin_2016,deleon_2013} and
empirical modeling \citep{barbour_2011,morris_2011,johnston_2015}. The
latter approeach models changes in biomass in response to changes in
mortality (\emph{i.e.} culling). In this case, biomass represents the
sum of all fish's individual weight. The individual weight of an
organism (Total Weight; TW) can be estimated from its Total Length (TL)
using the allometric growth equation (\(TW = aTL^b\)). Parameters \(a\)
and \(b\) for this equation exist for North Carolina
\citep{barbour_2011}, Northern \citep{fogg_2013,dahl_2014} and Southern
Gulf of Mexico \citep{aguilarperera_2016}, the Southern Mexican
Caribbean \citep{sabidoitza_2016}, Bahamas \citep{darling_2011}, Little
Cayman \citep{edwards_2014}, Jamaica \citep{chin_2016}, Bonaire
\citep{deleon_2013} Puerto Rico \citep{toledohernndez_2014}, and Costa
Rica \citep{sandel_2015}, but remain unavailable for the central Mexican
Caribbean. The weight-at-length of a species can vary across regions as
a response to biotic (\emph{e.g.} local food availability) and abiotic
(\emph{e.g.} water temperature) conditions \citep{johnson_2016}. Thus,
when using biomass-informed models or estimating weights from length
observations, it is important to use site-specific parameters. This is
especially important when research involves identifying the total
biomass available for harvest by fishers \citep{chin_2016} or the
efficacy of lionfish removals
\citep{barbour_2011,morris_2011,johnston_2015}.

This study reviews lionfish allometric growth throughout the invasion
range, and provides a new pair of parameters specific to the central
Meixcan Caribbean. The results suggest there are important
regional-scale variations in allometric growth patterns of lionfish. The
observed differences highlight the importance of using site-specific
parameters, especially when informing invasion management strategies.

The main objective of this work was to compare allometric growth of
lionfish throughout the invasion range. Allometric parameters were
retreived from scientific literature, and an additional pair of
parameters was calculated from filed observations in the central Mexican
Caribbean.

Length-weight relationships (n = 15) identified in literature were
obtained for North Carolina (n = 1, \citet{barbour_2011}), Northern (n =
3, \citet{fogg_2013}) and Southern Gulf of Mexico (n = 3,
\citet{aguilarperera_2016}), the Southern Mexican Caribbean (n = 1,
\citet{sabidoitza_2016}), Little Cayman (n = 1,\citet{edwards_2014}),
Jamaica (n = 1, \citet{chin_2016}), Bonaire (n = 1, \citet{deleon_2013})
and Costa Rica (n = 1, \citet{sandel_2015}). Locations with allometric
parameters are shown in Figure \ref{fig:map}. When available,
information on sampling methods, gender differentiation, location, and
depth ranges of each study was retrieved. Whenever gender was not
specified, it was assumed that the results were presented for pooled
genders.

Parameters from the central Mexican Caribbean were obtaines from data
collected in 10 sampling sites along the Mexican Caribbean coast in 2010
(Table S1). Sampling locations included wall and carpet reefs at depths
between 5.7 m and 38.1 m. All observed lionfish (n = 109) were collected
using hand nets and numbered collection bottles. The use of hand nets
prevented any weight loss due to bleeding and allowed a better
representation of small sizes by eliminating gear selectivity. Organisms
were euthanized (via pithing) and Total Length (TL; mm) and Total Weight
(TW; gr) were recorded before freezing organisms.

@@

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\label{fig:map}Locations where
allometric growth of lionfish (\emph{Pterois spp}) have been reported.
Sizes indicate sample size of each study, colors indicate the b
coefficient in Eq \ref{eq:allometric}}
\end{figure}

The weight at length relationship for lionfish in the central Mexican
Caribbean was calculated with the allometric growth function:

@@

Where \(Y = log_{10}(TW)\), \(X = log_{10}(TL)\), \(m = b\), and
\(c = log_{10}(a)\). Since \(b = m\), we will only use \(b\) throughout
the paper for simplicity. The coefficients (\(c\) and \(b\)) were
estimated with an Ordinary Least Square Regression and
heteroskedastic-robust standard error correction \citep{zeileis_2004}.
Both coefficients were tested against the null hypothesis of no change
(\emph{i.e.} \(H_0: c = 0\) and \(H_0: b = 0\)). Additionally, the
allometric parameter was tested against the null hypothesis of isometric
growth (\(H_0: b = 3\)). Coefficients were tested with a two-tailed
Student's t-test. The significance of the regression was corroborated
with an F-test.

During the review process, some studies indistinctly used \(a\) to
report either the ponderal index in \ref{eq:allometric} or the
y-intercept (\(c\)) in \ref{eq:log-alo-trans}. Others reported their
parameters as mm-to-gr conversions, but a rapid evaluation of such
parameters indicated that they were estimated as cm-to-gr conversions.
Here, all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies were transformed for consistency.

Since uncertainty around estimated relationships was not reported in
some of the reviewed studies, it was not possible to test for
statistical differences between relationships. Instead, the 16
length-weight relationships were used to calculate expected weight
length observations of the organisms sampled from the Central Mexican
Caribbean (n = 109). Expected weights were divided by the observed
weights to obtain a ratio. Difference in mean weight ratios across
studies were tested with a one-way Analysis of Variance (ANOVA).

All hypothesis testing was performed with an \emph{a priori} confidence
level of \(\alpha = 0.01\) in R version 3.4.0 \citep{rcore_2017}. Raw
data and code used in this work is available at dryad.org.

@@

The model adjusted to \ref{eq:log-alo-trans} estimated the coefficient
values at \(b = 3.2347391\) and \(c = -5.4940866\). Thus, TW (gr) can be
calculated from TL (mm) as a linear equation:
\(log_{10}(TW) = 3.2347391\times log_{10}(TL) -5.4940866\), or its
exponential form: \(TW = 3.2056297\times 10^{-6}\times TL^{3.2347391}\).
The intercept (\(c\)) and slope \((b)\) were significantly different
from zero (\(t(107) = -66.17; p<0.001\) and \(t(107) = 83.24; p<0.001\),
respectively), rejecting the null hypothesis of no change. Additionally,
the allometric factor (\(b\)) was significantly different from the value
of isometric growth of \(b = 3\) (\(t(107) = 6.04; p<0.001\)),
indicating that lionfish present allometric growth. More information on
model fit and confidence intervals for the estimated coefficients is
presented in TableS2. The relationship between Total Length and Total
Weight is presented in Figure \ref{fig:l-w-carib}.

From the 11 peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and the additional calculated for the
central Mexican Caribbean, 16 parameters were identified (Table
\ref{tab:all_params}, Fig \ref{fig:all_allo}). Two studies
\citep{aguilarperera_2016,fogg_2013} reported gender-level and pooled
parameters, while the rest presented pooled results. The smallest
coefficient of determination was presented by \citet{chin_2016} with
\(R^2 = 0.8715\), while \citet{sabidoitza_2016} reported the highest
value at \(R^2 = 0.9907\). Reviewed studies presented information for
organisms obtained at depths between 0.5 and 57 m. Three studies
\citep{aguilarperera_2016,chin_2016,dahl_2014} explicitly stated that
their organisms were sampled with pole spears. Five studies
\citep{sandel_2015,barbour_2011,fogg_2013,edwards_2014,sabidoitza_2016,toledohernndez_2014}
mentioned that some of their organisms were obtained with pole spears
(or other type of harpoon), but also hand-held nets or fish traps. Two
studies \citep{deleon_2013,darling_2011} did not specify how organisms
were sampled.

Parameters from models fit to males or females exclusively tend to have
a higher steepness (\emph{i.e.} higher allometric parameter), with mean
\(\pm\) standard deviation values of \(b = 3.27 \pm 0.06\) and
\(b = 3.31 \pm 0.23\) for males and females respectively, compared to
parameters from models for pooled genders with a mean \(\pm\) standard
deviation value of \(b = 3.13 \pm 0.22\). In the case of the ponderal
index (\(a\)) and its \(log_{10}\) transformation (\(c\)), values were
higher for parameters for pooled genders. Figure \ref{fig:all_allo}
shows the weight-at-length relationships with parameters from all
studies.

There were significant differences in expected-to-observed weight ratios
estimated for each pair of parameters (F(15, 1728) = 38.26; p
\textless{} 0.001). From all allometric parameters reviewed, those of
\citet{edwards_2014} provided the lowest weight estimates, with an
expected-to-observed weight ratio of 0.98 \(\pm\) 0.23 (mean \(\pm\)
SD). On the other hand,\citet{barbour_2011} yielded the highest weight
estimates, with a mean (\(\pm\) SD) expected-to-observed weight ratio of
1.76 \(\pm\) 0.50. Predicted-to-observed weight ratios and groups
identified by Tukey's HSD (\(\alpha = 0.05\)) are presented in Figure
\ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\label{fig:l-w-carib}Length-weight
relationship for 109 lionfish sampled in the Central Mexican Caribbean.
Points indicate samples, solid line indicates curve of best fit (See
Table S1).}
\end{figure}

\caption{\label{tab:unnamed-chunk-5}\label{tab:all_params}Summary of 13 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, a = scaling parameter for eq. 1, c = y-intercept for eq. 3, b = exponent or slope for eq. 1 or eq. 3, respectively. The $R^2$ column indicates reported model fit.}
\centering
\begin{tabular}[t]{lllrrll}
\toprule
Region & Sex & n & a & b & R2 & Reference\\
\midrule
Alacranes Reef, Mexico & B & 472 & 2.90e-06 & 3.30 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
Alacranes Reef, Mexico & F & 67 & 1.20e-06 & 3.47 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
Alacranes Reef, Mexico & M & 59 & 4.20e-06 & 3.23 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
Bahamas & B & - & 2.50e-06 & 3.29 & - & Darling et al., 2011\\
Bonaire & B & 1450 & 2.28e-05 & 2.89 & 0.96 & de Leon et al., 2013\\
\addlinespace
Costa Rica & B & 458 & 3.64e-05 & 2.81 & - & Sandel et al., 2015\\
Discovery Bay, Jamaica & B & 419 & 2.75e-05 & 2.85 & 0.8715 & Chin et al 2016\\
Little Cayman & B & 1887 & 3.00e-06 & 3.24 & 0.97 & Edwards et al., 2014\\
North Carolina & B & 774 & 2.89e-05 & 2.89 & - & Brbour et al 2011\\
Northern Gulf of Mexico & B & 934 & 2.10e-06 & 3.34 & 0.98 & Dahl \& Patterson, 2014\\
\addlinespace
Northern Gulf of Mexico & B & 582 & 1.40e-06 & 3.43 & 0.99 & Fogg et al., 2013\\
Northern Gulf of Mexico & M & 119 & 2.70e-06 & 3.31 & 0.97 & Fogg et al., 2013\\
Northern Gulf of Mexico & F & 115 & 6.80e-06 & 3.14 & 0.94 & Fogg et al., 2013\\
Puerto Aventuras, Mexico & B & 109 & 3.20e-06 & 3.23 & 0.9766 & This study\\
Puerto Rico & B & 227 & 8.00e-06 & 3.11 & 0.958 & Toledo-Hernández et al., 2014\\
Xcalak, Mexico & B & 2143 & 5.20e-06 & 3.18 & 0.9907 & Sabido-Itza et al., 2016\\
\bottomrule
\end{tabular}
\end{table}
